HISTO1

### 2022-11-25 code/Histology all rounds.Rmd
Progress on Imac

## Load data
```{r data load}
data <- read_excel("data/raw/Histo_analysis .xlsx")

allrounds <- read_excel("data/processed/weight_development.xlsx")

@@

histo <- filter_joined_df %>% 
  inner_join(clean_data, by = c("sample_id" = "ID"))

histo <- histo %>% 
  pivot_longer(., `PAS-RELATIVE-TO-TOTAL(%)`:`AB-RELATIVE-TO-TOTAL(%)`, 
               names_to = "AB-PAS",
               values_to = "values")

@@

```{r}
histo %>%  ### Try to wrap histo longer to have all stain values in 1 column and facet-wrap by PAS + AB
  filter(TISSUE == "ileum") %>%
  filter('AB-PAS' == "PAS-RELATIVE-TO-TOTAL(%)") %>% 
  ggplot(
    aes(
    x = group2, 
    y = values,
    fill = group2)
    ) +
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(~ exp_number,
             scales = "free") +
  labs(
    #title =  "Ileum AB-PAS staining % of total tissue", subtitle = "Faceted by experimental round (2-4)",
       y = "PAS stained as % of total tissue",
       x = "Maternal Treatment") + 
  theme_bw() +
  theme(
    axis.text.x = element_blank()
  )

@@

@@

>>>>

## Load data
```{r data load}
data <- read_excel("data/raw/Histo_analysis .xlsx") %>% 
  rename(pas_pct = `PAS-RELATIVE-TO-TOTAL(%)`,
         ab_pct = `AB-RELATIVE-TO-TOTAL(%)`) %>% 
  mutate(AB_PAS_pct = (ab_pct +pas_pct)/2) %>% 
  mutate(tissue = if_else(TISSUE == "colon","Colon","Ileum"))

allrounds <- read_excel("data/processed/weight_development.xlsx")  %>% 
  select(-c(cage:comments))

@@

histo <- filter_joined_df %>% 
  inner_join(clean_data, by = c("sample_id" = "ID")) %>% 
  rename(Group = group2)
```

# Exp2 
```{r Exp2-ileum}
Exp2_plot_histo <- histo %>% 
  #filter(TISSUE == "ileum") %>%
  filter(exp_number == 2) %>% 
  ggplot(
    aes(
    x = Group, 
    y = AB_PAS_pct,
    fill = Group)
    ) +
  facet_wrap(~tissue,
             scales = "free")+
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  #coord_cartesian(ylim = c(0, 10)) +
  stat_compare_means(aes(label = ..p.signif.., ),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
       y = "Mucin (% of total tissue)",
       x = "") + 
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1") + plot_annotation (
  title = "Image analytical quantification of mycin",
   #subtitle = 'All cytokines measured',
  #tag_levels = 'A',
  #caption = 'CON = Water, BF = Breast-Feeding, GENTA = Gentamicin, VANCO = Vancomycin,',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold")))

Exp2_plot_histo
 ggsave(width = 110, height = 110, units = "mm", 'exports/histo/Exp2-ab_pas_plot.png',Exp2_plot_histo)

@@

# Exp3 
```{r Exp3-ileum}
Exp3_plot_histo <- histo %>% 
  filter(Group != "VANCO-BF") %>%
  filter(exp_number == 3) %>% 
  ggplot(
    aes(
    x = Group, 
    y = AB_PAS_pct,
    fill = Group)
    ) +
  facet_wrap(~tissue,
             scales = "free")+
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  #coord_cartesian(ylim = c(0, 10)) +
  stat_compare_means(aes(label = ..p.signif.., ),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
       y = "Mucin (% of total tissue)",
       x = "") + 
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1") + plot_annotation (
  title = "Image analytical quantification of mycin",
   #subtitle = 'All cytokines measured',
  #tag_levels = 'A',
  #caption = 'CON = Water, BF = Breast-Feeding, GENTA = Gentamicin, VANCO = Vancomycin,',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold")))

Exp3_plot_histo
 ggsave(width = 170, height = 110, units = "mm", 'exports/histo/Exp3-ab_pas_plot.png',Exp3_plot_histo)

@@

# Exp4 
```{r Exp4-ileum}
Exp4_plot_histo <- histo %>% 
  #filter(TISSUE == "ileum") %>%
  filter(exp_number == 4) %>% 
  ggplot(
    aes(
    x = Group, 
    y = AB_PAS_pct,
    fill = Group)
    ) +
  facet_wrap(~tissue,
             scales = "free")+
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  #coord_cartesian(ylim = c(0, 10)) +
  stat_compare_means(aes(label = ..p.signif.., ),
                   method = "wilcox.test", ref.group = "VANCO-SM",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
       y = "Mucin (% of total tissue)",
       x = "") + 
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1") + plot_annotation (
  title = "Image analytical quantification of mycin",
   #subtitle = 'All cytokines measured',
  #tag_levels = 'A',
  #caption = 'CON = Water, BF = Breast-Feeding, GENTA = Gentamicin, VANCO = Vancomycin,',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold")))

Exp4_plot_histo
 ggsave(width = 110, height = 110, units = "mm", 'exports/histo/Exp4-ab_pas_plot.png',Exp4_plot_histo)

@@

# Overall MAT
```{r overall-mat}
histo_overall_colon <- histo %>% 
  filter(tissue == "Colon") %>%
  filter(exp_number != 4) %>% 
  ggplot(
    aes(
    x = received_antibiotics, 
    y = AB_PAS_pct,
    fill = received_antibiotics)
    ) +
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  facet_wrap(~exp_number,
             scales = "free") +
  #coord_cartesian(ylim = c(0, 40)) +
  stat_compare_means(aes(label = paste0("p = ", ..p.format..)),
                   method = "wilcox.test", ref.group = "No",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
      y = "Mucin (% of total tissue)",
       x = "",
       fill='') + 
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_blank()) +  
  #legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1")

histo_overall_colon

histo_overall_ileum <- histo %>% 
  filter(tissue == "Ileum") %>%
  filter(exp_number != 4) %>% 
  ggplot(
    aes(
    x = received_antibiotics, 
    y = AB_PAS_pct,
    fill = received_antibiotics)
    ) +
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  facet_wrap(~exp_number,
             scales = "free") +
  #coord_cartesian(ylim = c(0, 40)) +
  stat_compare_means(aes(label = paste0("p = ", ..p.format..)),
                   method = "wilcox.test", ref.group = "No",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
      y = "Mucin (% of total tissue)",
       x = "",
      fill='Maternal Antibiotic Treatment') +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1")

histo_overall_ileum

histo_overall_both <- (histo_overall_colon/histo_overall_ileum) + plot_annotation (
  title = "Effect of Maternal Antibiotics Treatment on intestinal mucin",
   #subtitle = 'All cytokines measured',
  tag_levels = 'A',
  #caption = 'CON = Water, BF = Breast-Feeding, GENTA = Gentamicin, VANCO = Vancomycin,',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold")))

histo_overall_both
 ggsave(width = 110, height = 110, units = "mm", 'exports/histo/Hist_MAT',histo_overall_both)
 ggsave('exports/histo/Hist_MAT.png', histo_overall_both)

# Overall FEED
```{r overall-mat}
histo_FEED_colon <- histo %>% 
  filter(tissue == "Colon") %>%
  filter(exp_number != 2) %>% 
  ggplot(
    aes(
    x = type_of_feed, 
    y = AB_PAS_pct,
    fill = type_of_feed)
    ) +
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  facet_wrap(~exp_number,
             scales = "free") +
  #coord_cartesian(ylim = c(0, 40)) +
  stat_compare_means(aes(label = paste0("p = ", ..p.format..)),
                   method = "wilcox.test", ref.group = "BF",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
      y = "Mucin (% of total tissue)",
       x = "",
       fill='') + 
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_blank()) +  
  #legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1")

histo_FEED_colon

histo_FEED_ileum <- histo %>% 
  filter(tissue == "Ileum") %>%
  filter(exp_number != 2) %>% 
  ggplot(
    aes(
    x = type_of_feed, 
    y = AB_PAS_pct,
    fill = type_of_feed)
    ) +
  geom_boxplot(outlier.shape = NA) +
  geom_point() +
  facet_wrap(~exp_number,
             scales = "free") +
  #coord_cartesian(ylim = c(0, 40)) +
  stat_compare_means(aes(label = paste0("p = ", ..p.format..)),
                   method = "wilcox.test", ref.group = "BF",
                   label.y.npc = 0.85) +
  labs(
    #title =  ,
      y = "Mucin (% of total tissue)",
       x = "",
      fill='Maternal Antibiotic Treatment') +
  theme_bw()+
  theme(                   # Change the appearance of labels
    plot.title = element_text(color="darkgrey", size=12, face="bold"),
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    #axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.title = element_text(face="bold"),   
    legend.position="bottom",
    legend.background = element_rect(fill="white")) + # Change legend box background color
  scale_fill_brewer(palette="Set1")

histo_FEED_ileum

histo_FEED_both <- (histo_FEED_colon/histo_FEED_ileum) + plot_annotation (
  title = "Effect of formula or breastmilk on intestinal mucin",
   #subtitle = 'All cytokines measured',
  tag_levels = 'A',
  #caption = 'CON = Water, BF = Breast-Feeding, GENTA = Gentamicin, VANCO = Vancomycin,',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold")))

histo_FEED_both
 ggsave(width = 110, height = 110, units = "mm", 'exports/histo/Hist_MAT',histo_overall_both)
 ggsave('exports/histo/Hist_FEED.png', histo_FEED_both)
